Abstract Data Type Inference

### 2015-03-18 histories.tex

In this work, histories are abstractions of executions which retain method
names yet lose the exact argument and return values, and retain the relative
order of operations yet lose the exact sequence of call and return actions.
Formally, a \emph{history} is a tuple $h = \tup{O,<,c,f,m}$ where
\begin{itemize}

@@

  \item $f: O \to \mathbb{M}$ labels operations with method names, and

  \item $m: O \rightharpoonup O$ is a partial \emph{matching} function.

@@

Two operations $o_1, o_2 \in O$ are \emph{identical} when they have
\begin{itemize}

  \item the same labels: $c(o_1) = c(o_2)$, $f(o_1)= f(o_2)$, and

  \item the same matching: either $m(o_1) = m(o_2)$, or both $m(o_1)$
  and $m(o_2)$ are undefined.

@@

Our abstraction of executions relies on correlating the operations associated
with the same values in an execution via the partial matching function of a
history. We construct partial matching functions systematically. A
\emph{matching scheme} $M$ associates to each execution $e$ with operations $O$
a partial matching function $M(e): O \rightharpoonup O$.

@@

The history $H(e,M)$ of an execution $e$ under matching scheme $M$ is the tuple
$\tup{O,<,c,f,M(e)}$ where
\begin{itemize}

@@

\end{itemize}
We denote the set $\set{H(e,M) : e \in E}$ of histories of an execution set $E$
by $H(E,M)$.

@@

A matching scheme $M$ is \emph{faithful} to a set $E$ of executions when $e \in
E$ if{f} $e' \in E$ for any two executions $e$ and $e'$ such that $H(e,M) =
H(e',M)$. A set of executions $E$ is \emph{data independent}\footnote{TODO NOTE
THAT THIS IS NOT THE STANDARD DEFINITION…} when there exists a faithful
matching scheme. By definition, our abstraction of executions incurs no loss of
precision for data-independent execution sets.

@@

  $H(e,M) \in H(E,M)$ if and only if $e \in E$, for any faithful matching
  scheme $M$.

@@

In Section~\ref{sec:nature} we demonstrate faithful matching schemes for the
executions of naturally-occurring ADTs, and in Section~\ref{sec:matching} we
demonstrate how to infer faithful matching schemes. Otherwise, for the
remainder of this work, we assume each set of executions comes equipped with a
faithful matching scheme $M$, and abbreviate $H(E,M)$ by $H(E)$.

>>>>

In this work, histories are abstractions of executions which retain method
names yet lose the exact argument and return values, and retain the relative
order of operations yet lose the exact sequence of call and return actions.
Formally, a \emph{history} is a tuple $h = \tup{O,<,c,f,m,r}$ where
\begin{itemize}

@@

  \item $f: O \to \mathbb{M}$ labels operations with method names,

  \item $m: O \rightharpoonup O$ is a partial \emph{matching} function, and

  \item $r: O \to \mathbb{B}$ labels operations as \emph{read-only}, or not.

@@

Two operations $o_1, o_2 \in O$ are \emph{identical} when
\begin{itemize}

  \item they have the same labels: $c(o_1) = c(o_2)$, $f(o_1)= f(o_2)$, and
  $r(o_1) = r(o_2)$, and

  \item they have the same matching: either $m(o_1) = m(o_2)$, or both $m(o_1)$
  and $m(o_2)$ are undefined.

@@

Our abstraction of executions relies on correlating the operations associated
with the same values in an execution via the partial matching function of a
history. We construct partial matching functions systematically. A
\emph{matching scheme} $\tup{M,R}$ associates to each execution $e$ with
operations $O$ a partial matching function $M(e): O \rightharpoonup O$ and a
read-only operation labeling $R(e): O \to \mathbb{B}$.

@@

The history $H(e,M,R)$ of an execution $e$ under matching scheme $\tup{M,R}$ is
the tuple $\tup{O,<,c,f,M(e),R(e)}$ where
\begin{itemize}

@@

\end{itemize}
We denote the set $\set{H(e,M,R) : e \in E}$ of histories of an execution set
$E$ by $H(E,M,R)$.

@@

A matching scheme $\tup{M,R}$ is \emph{faithful} to a set $E$ of executions
when $e \in E$ if{f} $e' \in E$ for any two executions $e$ and $e'$ such that
$H(e,M,R) = H(e',M,R)$. A set of executions $E$ is \emph{data
independent}\footnote{TODO NOTE THAT THIS IS NOT THE STANDARD DEFINITION…} when
there exists a faithful matching scheme. By definition, our abstraction of
executions into histories incurs no loss of precision for data-independent
execution sets.

@@

  $H(e,M,R) \in H(E,M,R)$ if and only if $e \in E$, for any faithful matching
  scheme $\tup{M,R}$.

@@

In Section~\ref{sec:nature} we demonstrate faithful matching schemes for the
executions of naturally-occurring ADTs, and in Section~\ref{sec:matching} we
demonstrate how to infer faithful matching schemes. Otherwise, for the
remainder of this work, we assume each set of executions comes equipped with a
faithful matching scheme $\tup{M,R}$, and abbreviate $H(E,M,R)$ by $H(E)$.
